---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dfvad

<!-- badges: start -->
<!-- badges: end -->

## Overview

`dfvad` decomposes value added growth into explanatory factors.
A cost constrained value added function is defined to specify the 
production frontier. Industry estimates can also be aggregated using 
a weighted average approach.

## Installation

`dfvad` is available from https://github.com/shipei-zeng/dfvad. To install it, `install_github` from the `devtools` package can be helpful.

``` {r eval=FALSE}
devtools::install_github("shipei-zeng/dfvad")
```

If error messages show that the URL cannot be opened, please set the download option before installing it.

``` {r eval=FALSE}
options(download.file.method = "libcurl")
```

If error messages show that schannel failed to receive handshake, please delete the previous package before updating it.

It has also been uploaded to the CRAN repository, which can be downloaded using:

``` {r eval=FALSE}
installed.packages("dfvad")
```

## Usage

`value_decom()` for decomposing nominal value added growth identifies the contributions from efficiency change, growth of primary inputs, changes in output and input prices, technical progress and returns to scale.

```{r}
library(dfvad)
# Use the built-in dataset "mining"
table1 <- value_decom(c("h2","x2"), c("w2","u2"), "y2", "p2", "year", mining)[[1]]
head(table1)
table2 <- value_decom(c("h2","x2"), c("w2","u2"), "y2", "p2", "year", mining)[[2]]
head(table2)
```

`t_weight()` follows a "bottom up" approach that uses weighted averages of the sectoral decompositions to provide an approximate decomposition into explanatory components at the aggregate level.

```{r}
library(dfvad)
# Use the built-in dataset "sector"
table1 <- t_weight("y", "p", "industry", "year", "alpha", "beta", "gamma", "epsilon", "tau", sector)[[1]]
head(table1)
table2 <- t_weight("y", "p", "industry", "year", "alpha", "beta", "gamma", "epsilon", "tau", sector)[[2]]
head(table2)
```

## Display

Here is an example to plot the explanatory factors of productivity (logarithmic indexes). Additional packages such as `ggplot2` and `reshape2` are required.

```{r out.width = "60%", dpi=600}
library(dfvad)
library(ggplot2)
library(reshape2)
# Get the decomposition result
df <- value_decom(c("h2","x2"), c("w2","u2"), "y2", "p2", "year", mining)[[2]]
# Extract columns and rename
df_cmpt <- data.frame(df[,"period"], log(df[,c("T", "E", "C")]))
colnames(df_cmpt) <- c("year", "lnT", "lnE", "lnC")
df_tfp <- data.frame(df[,"period"], log(df[,"TFP"]))
colnames(df_tfp) <- c("year", "lnTFP")
# Set the colour scheme
palette_a <- c("goldenrod1", "seashell4", "red")
# Convert data into a tidy form
df_cmpt_tidy <- melt(df_cmpt, id.vars="year")
# Plot the components
plot_out <- ggplot(df_cmpt_tidy) + geom_bar(aes(x=year, y=value, fill=variable), stat="identity") +
        geom_line(data=df_tfp, aes(x=year,y=lnTFP,color='black'), lwd=0.5) +
        ylab('Log Index') + xlab('Year') + 
        scale_fill_manual("", values=palette_a) + 
        scale_colour_manual("", values=c('black'='black'), labels = c('lnTFP')) + 
        scale_x_continuous(breaks = seq(min(df$period), max(df$period), by = 3)) + 
        theme_classic()
print(plot_out)
```
